---
title: "Case Study 2: Employee Attrition"
author: "Catherine Ticzon" 
output: html_document
editor_options: 
  chunk_output_type: console
---

## Summary  

This Rmarkdown file contains the supporting code for DDS Case Study 2 on employee attrition.  The file is organized in chunks as follows:  
  
  i.  Set Up  
  1.  Data Cleaning  
  2.  Exploratory Data Analysis (EDA)  
  3.  Predicting Attrition with knn   
  4.  Predicting Attrition with naive bayes  
  5.  Predicting Attrition with linear regression  
  6.  Predicting Salary with linear regression  

## i. Set Up 
```{r setup, warning=FALSE}
library(RCurl)
library(aws.s3)
library(tidyverse)
library(dplyr)
library(ggplot2)

Sys.setenv("AWS_ACCESS_KEY_ID" = "############",
           "AWS_SECRET_ACCESS_KEY" = "############",
           "AWS_DEFAULT_REGION" = "us-east-2")

# Need AWS_DEFAULT_REGION or else Error when loading in data from s3:
# Error in parse_aws_s3_response(r, Sig, verbose = verbose) : 
# Moved Permanently (HTTP 301).

data <- s3read_using(FUN = read.csv,
                     bucket = "ddsproject1",
                     object = "CaseStudy2-data.csv")
```

## 1. Data Cleaning   

  * In this section, I make sure I do not have any missing values and I also change some of the variables to factor levels.  
  
```{r, results=FALSE, warning=FALSE}
sum(is.na(data)) # No missing data 

# Check variables and change character values in columns to factor levels 

table(data$Attrition)
data$Attrition = factor(data$Attrition, levels = c("No", "Yes"))

table(data$BusinessTravel)
data$BusinessTravel = factor(data$BusinessTravel, levels = c("Non-Travel", "Travel_Frequently",
                                                             "Travel_Rarely"))

table(data$Department) 
data$Department = factor(data$Department, levels = c("Human Resources", "Research & Development", "Sales"))

table(data$EducationField)
data$EducationField = factor(data$EducationField, levels = c("Human Resources", "Life Sciences", "Marketing", "Medical", "Other", "Technical Degree"))

table(data$Gender)
data$Gender = factor(data$Gender, levels = c("Female", "Male"))

table(data$JobRole)
data$JobRole = factor(data$JobRole, levels = c("Healthcare Representative", "Human Resources", "Laboratory Technician", "Manager", "Manufacturing Director", "Research Director", "Research Scientist", "Sales Executive", "Sales Representative"))

table(data$MaritalStatus)
data$MaritalStatus = factor(data$MaritalStatus, levels = c("Divorced", "Married", "Single"))

```

## 2. Exploratory Data Analysis (EDA)    

  * In this section, I conduct and exploratory data analysis (EDA) to find trends in employee attrition   
  * I also conduct an EDA on the entire sample to compare to the trends in attrition in order to find variables that disproportionately make-up employee attrition  
  * This section includes several plots that are hidden in the knitted file  
  
```{r, warning=FALSE, fig.show='hide'}
# install.packages("ggthemes")
library(ggthemes)

# Explore variables of employees who left

# Create variable of those who left
quit <- data %>%
  filter(Attrition == "Yes")

# Summarize attrition by Gender
ggplot(quit, mapping=aes(x=Gender)) +
  geom_bar() +
  geom_text(stat='count', aes(label= ..count..), vjust= -1) + 
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank())

# Summarize attrition by Age
attr_Age <- quit %>%
  group_by(Age) %>%
  summarise(count = n())

ggplot(attr_Age, mapping=aes(x=Age, y=count)) +
  geom_col()

ggplot(attr_Age, mapping=aes(x=Age)) +
  geom_boxplot()


# Summarize attrition by Job Role ----------------------------------------------
attr_JobRole <- data %>%
  group_by(JobRole, Attrition) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = Attrition, values_from = count, values_fill = 0) 

attr_JobRole_perc <- attr_JobRole %>%
  mutate(Yes_Perc = (Yes /140) * 100) 

ggplot(attr_JobRole_perc, mapping=aes(x=JobRole, y=Yes_Perc)) +
  geom_col(fill="black", width= 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", Yes_Perc), hjust=-.5)) + 
    theme(axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
  coord_flip() 


# Summarize attrition by Travel 
attr_Travel <- data %>%
  group_by(BusinessTravel, Attrition) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = Attrition, values_from = count, values_fill = 0) 


# Summarize attrition by DailyRate
attr_DailyRate <- data %>%
  group_by(DailyRate, Attrition) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = Attrition, values_from = count, values_fill = 0) 

  
# Create classes for DailyRate: Lower (100-500), Middle (500-1000), Upper (1000-1500)
breaks_DailyRate <- data %>% 
  mutate(DailyRateClass = cut(data$DailyRate, breaks = c(100,500,1000,1500), labels = c("Lower", "Middle", "Upper")))
                        
attr_breaks_DailyRate <- breaks_DailyRate %>%
  group_by(DailyRateClass, Attrition) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = Attrition, values_from = count, values_fill = 0) 

ggplot(data, mapping = aes(x=DailyRate, y=Attrition)) +
  geom_boxplot()

 
# Summarize attrition by Department --------------------------------------------
attr_Department <- quit %>%
  group_by(Department) %>%
  summarise(count = n())

attr_Department_perc <- attr_Department %>%
  mutate(Department_perc = (count / 140) * 100)

ggplot(attr_Department_perc, mapping=aes(x=Department, y=Department_perc)) +
  geom_col(fill="black", width = 0.4) +
  geom_text(aes(label = sprintf("%.1f%%", Department_perc), vjust=2, color="white")) + 
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
   scale_color_identity()


# Plot Attrition by Department next to whole sample

#prop_department <- full_join(attr_Department_perc, Department_perc, by="Department") %>%
#  rename("AttritionPerc"= "Department_perc",
#         "SamplePerc" = "perc")


#ggplot(prop_department, aes(x = Department)) +
  #geom_bar(aes(y = AttritionPerc, fill = "AttritionPerc"), position = "dodge", stat = "identity") +
  #geom_bar(aes(y = SamplePerc, fill = "SamplePerc"), position = "dodge", stat = "identity") +
  #scale_fill_manual(values = c("AttritionPerc" = "red", "SamplePerc" = "blue")) +
  #labs(x = "Department", y = "Percentage")


# Summarize attrition by DistanceFromHome
attr_Distance <- quit %>%
  group_by(DistanceFromHome) %>%
  summarise(count = n())

ggplot(attr_Distance, mapping=aes(x=DistanceFromHome, y=count)) +
  geom_col(fill="black") +
  labs(x = "Distance from Home (Units)", y = "Number of Employees") +
  #geom_text(aes(label = count), vjust = -0.3, color = "black") + 
  theme(axis.title.y = element_blank(),
        #axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        #axis.text.y = element_blank(),
        #axis.ticks = element_blank(),
        panel.grid.major.y = element_line(color="light grey"),
        #panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) 

# Summarize attrition by Education
attr_Education <- quit %>%
  group_by(Education) %>%
  summarise(count = n())

ggplot(attr_Education, mapping=aes(x=Education, y=count)) +
  geom_col(fill="light gray") +
  geom_text(aes(label = count), vjust = -0.3, color = "black") + 
  labs(x="Education Level",
       y= "Number of Employees") +
  theme(axis.title.y = element_blank(),
        #axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        #panel.grid.major.y = element_line(color="light grey"),
        #panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank())

# Summarize attrition by EducationField 

attr_EducationField <- quit %>%
  group_by(EducationField) %>%
  summarise(count = n())

ggplot(attr_EducationField, mapping=aes(x=EducationField, y=count)) +
  geom_col()

# Attriton by JobLevel----------------------------------------------------------
attr_JobLevel <- quit %>%
  group_by(JobLevel) %>%
  summarise(count = n())

ggplot(quit, mapping=aes(x=JobLevel)) +
  geom_bar()

attr_JobLevel_perc <- attr_JobLevel %>%
  mutate(JobLevel_perc = (count / 140) * 100)

ggplot(attr_JobLevel_perc, mapping=aes(x=JobLevel, y=JobLevel_perc)) +
  geom_col(fill="light gray", width=.5) + 
  geom_text(aes(label = sprintf("%.1f%%", JobLevel_perc))) +
  theme(axis.title.y = element_blank(), 
        #axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
  labs(x= "Job Level")


# Attrition by EnvironmentSatisfaction------------------------------------------
attr_EnvSat <- quit %>%
  group_by(EnvironmentSatisfaction) %>%
  summarise(count=n()) %>%
  mutate(EnvSat_perc = (count/140) *100)

ggplot(attr_EnvSat, mapping=aes(x=EnvironmentSatisfaction, y=EnvSat_perc)) +
  geom_col(fill="light gray", width=.5) + 
  geom_text(aes(label = sprintf("%.1f%%", EnvSat_perc))) +
  theme(axis.title.y = element_blank(), 
        #axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
  labs(x="Environment Satisfaction")

# Attrition by WorkLifeBalance
attr_WorkLifeBalance <- quit %>%
  group_by(WorkLifeBalance) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_WorkLifeBalance, mapping=aes(x=WorkLifeBalance, y=perc)) +
  geom_col()

# Attrition by RelationshipSatisfaction
attr_RelationshipSat <- quit %>%
  group_by(RelationshipSatisfaction) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_RelationshipSat, mapping=aes(x=RelationshipSatisfaction, y=perc)) +
  geom_col()

# Attrition by JobSatisfaction
attr_JobSat <- quit %>%
  group_by(JobSatisfaction) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_JobSat, mapping=aes(x=JobSatisfaction, y=perc)) +
  geom_col()

# Attrition by PerformanceRating
attr_PerfRate <- quit %>%
  group_by(PerformanceRating) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_PerfRate, mapping=aes(x=PerformanceRating, y=perc)) +
  geom_col()

# Attrition by PercentSalaryHike
attr_PercSalHike <- quit %>%
  group_by(PercentSalaryHike) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_PercSalHike, mapping=aes(x=PercentSalaryHike, y=perc)) +
  geom_col()

# Attrition by JobInvolvement-------------------------------------------------- 
attr_JobInvolv <- quit %>%
  group_by(JobInvolvement) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_JobInvolv, mapping=aes(x=JobInvolvement, y=perc)) +
  geom_col()

# Attrition by YearsSinceLastPromotion 
attr_YrsPromo <- quit %>%
  group_by(YearsSinceLastPromotion) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_YrsPromo, mapping=aes(x=YearsSinceLastPromotion, y=perc)) +
  geom_col()

# Attrition by TrainingTimesLastYear 
attr_Training <- quit %>%
  group_by(TrainingTimesLastYear) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_Training, mapping=aes(x=TrainingTimesLastYear, y=perc)) +
  geom_col()

# Attrition by MonthlyIncome
attr_MonthInc <- quit %>%
  group_by(MonthlyIncome) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/140) *100)

ggplot(attr_MonthInc, mapping=aes(x=MonthlyIncome, y=perc)) +
  geom_jitter()

ggplot(attr_MonthInc, mapping=aes(x=MonthlyIncome, y=count)) +
  geom_jitter()

# Explore variables of entire sample =======================================

# Age
ggplot(data, mapping=aes(x=Age)) +
  geom_histogram()


# Gender
ggplot(data, mapping=aes(x=Gender, fill=Gender)) +
  geom_bar() +
  geom_text(stat='count', aes(label= ..count..), vjust= -1) +
  scale_fill_manual(values = c("#000000", "#c9c9c9")) +
  theme(axis.title.y = element_blank(), 
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank())


# Job Roles --------------------------------------------------------------------
JobRole <- data %>%
  group_by(JobRole) %>%
  summarise(count = n()) 

JobRole_perc <- JobRole %>%
  mutate(JobRolePerc = (count /870) * 100) 

# Department 
Department <- data %>%
  group_by(Department) %>%
  summarise(count = n()) 

Department_perc <- Department %>%
  mutate(perc = (count /870) * 100) 

ggplot(Department_perc, mapping=aes(x=Department, y=perc)) +
  geom_col(fill="light grey", width=.4) + 
  geom_text(aes(label = sprintf("%.1f%%", perc), vjust= 1.5)) +
  theme(axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) 

# Plot Job Role by Count
ggplot(data, mapping=aes(x=JobRole)) +
  geom_bar(fill = "light grey") +
  geom_text(stat='count', aes(label= ..count..), hjust= -1) +
  theme(axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
  coord_flip()

# Plot Job Role by Percentage
ggplot(JobRole_perc, mapping=aes(x=JobRole, y=JobRolePerc)) +
  geom_col(fill = "light grey", width= 0.5) +
  geom_text(aes(label = sprintf("%.1f%%", JobRolePerc), hjust=1.5)) + 
  theme(axis.title.y = element_blank(), 
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
  coord_flip()

# Education
ggplot(data, mapping=aes(Education)) +
  geom_bar()

# Education Field
ggplot(data, mapping=aes(EducationField)) +
  geom_bar() 

# Job Level---------------------------------------------------------------------
JobLevel <- data %>%
  group_by(JobLevel) %>%
  summarise(count = n()) %>%
  mutate(JobLevel_perc = (count/870)*100)

ggplot(JobLevel, mapping=aes(x=JobLevel, y=JobLevel_perc)) +
  geom_col(fill="light gray", width=.5) + 
  geom_text(aes(label = sprintf("%.1f%%", JobLevel_perc))) +
  theme(axis.title.y = element_blank(), 
        #axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
  labs(x= "Job Level")

# Environment Satisfaction------------------------------------------------------
EnvSat <- data %>%
  group_by(EnvironmentSatisfaction) %>%
  summarise(count=n()) %>%
  mutate(EnvSat_perc = (count/870) *100)

ggplot(EnvSat, mapping=aes(x=EnvironmentSatisfaction, y=EnvSat_perc)) +
  geom_col(fill="light gray", width=.5) + 
  geom_text(aes(label = sprintf("%.1f%%", EnvSat_perc))) +
  theme(axis.title.y = element_blank(), 
        #axis.title.x = element_blank(),
        #axis.text.x = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.background = element_blank()) +
  labs(x="Environment Satisfaction")

# WorkLifeBalance
WorkLifeBalance <- data %>%
  group_by(WorkLifeBalance) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(WorkLifeBalance, mapping=aes(x=WorkLifeBalance, y=perc)) +
  geom_col()

# RelationshipSatisfaction
RelationshipSat <- data %>%
  group_by(RelationshipSatisfaction) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(RelationshipSat, mapping=aes(x=RelationshipSatisfaction, y=perc)) +
  geom_col()

# JobSatisfaction
JobSat <- data %>%
  group_by(JobSatisfaction) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(JobSat, mapping=aes(x=JobSatisfaction, y=perc)) +
  geom_col()

# PerformanceRating
PerfRate <- data %>%
  group_by(PerformanceRating) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(PerfRate, mapping=aes(x=PerformanceRating, y=perc)) +
  geom_col()

# PercentSalaryHike
PercSalHike <- data %>%
  group_by(PercentSalaryHike) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(PercSalHike, mapping=aes(x=PercentSalaryHike, y=perc)) +
  geom_col()

# JobInvolvement ---------------------------------------------------------------
JobInvolv <- data %>%
  group_by(JobInvolvement) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(JobInvolv, mapping=aes(x=JobInvolvement, y=perc)) +
  geom_col()

# YearsSinceLastPromotion 
YrsPromo <- data %>%
  group_by(YearsSinceLastPromotion) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(YrsPromo, mapping=aes(x=YearsSinceLastPromotion, y=perc)) +
  geom_col()

# TrainingTimesLastYear 
Training <- data %>%
  group_by(TrainingTimesLastYear) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(Training, mapping=aes(x=TrainingTimesLastYear, y=perc)) +
  geom_col()

# MonthlyIncome
MonthInc <- data %>%
  group_by(MonthlyIncome) %>%
  summarise(count=n()) %>%
  mutate(perc = (count/870) *100)

ggplot(MonthInc, mapping=aes(x=MonthlyIncome, y=perc)) +
  geom_jitter()

ggplot(MonthInc, mapping=aes(x=MonthlyIncome, y=count)) +
  geom_jitter()


# Misc plotting and EDA ------------------------------------------------------


# group_by JobRole 

jr_income <- data %>%
  group_by(JobRole) %>%
  summarise(mean_income = mean(MonthlyIncome))

jr_jobinvolv <- data %>%
  group_by(JobRole) %>%
  summarise(mean_jobinvolv = mean(JobInvolvement))

jr_envsat <- data %>%
  group_by(JobRole) %>%
  summarise(mean_envsat = mean(EnvironmentSatisfaction))

# Boxplot of JobLevel by JobRole
ggplot(data, mapping=aes(x=JobRole, y=JobLevel)) +
  geom_boxplot() +
  labs(title="Boxplot of Job Level by Job Role",
       subtitle="n=870",
       y="Job Level") + 
  theme(axis.title.y = element_blank(), 
        #axis.title.x = element_blank(),
        #axis.text.x = element_text(angle=90, hjust =.95, vjust= .2),
        axis.ticks = element_blank(),
        panel.grid.major.y = element_line(color="grey"),
        panel.background = element_blank()) +
  coord_flip()


jr_promo <- data %>%
  group_by(JobRole) %>%
  summarise(mean_promo = mean(YearsSinceLastPromotion))

ggplot(data, aes(x=JobRole, y=YearsAtCompany)) +
  geom_boxplot() +
  coord_flip() +
  theme(axis.title.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color="grey"),
        panel.grid.minor.x = element_line(color="grey"),
        #panel.grid.minor.y = element_line(color="grey"),
        panel.background = element_blank()) +
  labs(title= "Boxplot of Number of Years at Company by Job Role",
       subtitle ="n=870",
        x="Job Role", y="Number of Years at Company")
```


## 3. Analysis of Variables Related to Attrition  
```{r}

# Model JobLevel ~ Attrition
model_JobLevel <- lm(JobLevel ~ Attrition, data = data)
summary(model_JobLevel)

ggplot(data, aes(x=JobLevel, y=Attrition)) +
  geom_jitter()

# Plot JobLevel, Job Role, and Attrition

data %>% ggplot(aes(x=JobRole, y = JobLevel, color=Attrition)) +
  geom_jitter(alpha = .5) +
  theme(axis.text.x = element_text(angle=90, hjust =.95, vjust= .2),
        panel.grid.major.y = element_line(color="grey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_line(color="grey"),
        panel.grid.minor.y = element_line(color="grey"),
        panel.background = element_blank()) +
  labs(title= "Attrition Cases by Job Role and Job Level",
       subtitle ="n=870",
        x="Job Role", y="Job Level") 


# Plot Environment Satisfaction, Job Level, Departmen ~ Attrition 

theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(data, aes(EnvironmentSatisfaction, JobLevel)) + 
  labs(subtitle="n=870",
       title="Attrition by Environment Satisfaction, Job Level, and Department",
       x="Environment Satisfaction",
       y="Job Level")

g + geom_jitter(aes(col=Attrition, shape = Department), size=3, alpha=.7) + 
  geom_smooth(aes(col=Attrition), method="lm", se=F) +
  scale_color_manual(values = c("No" = "grey", "Yes" = "red")) +
  theme(panel.grid.major.y = element_line(color="grey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_line(color="grey"),
        panel.grid.minor.y = element_line(color="grey"),
        panel.background = element_blank())

# Switch x and y axis 

theme_set(theme_bw())  # pre-set the bw theme.
h <- ggplot(data, aes(JobLevel, EnvironmentSatisfaction)) + 
  labs(subtitle="n=870",
       title="Attrition by Environment Satisfaction, Job Level, and Department",
       x="Job Level",
       y="Environment Satisfaction")

h + geom_jitter(aes(col=Attrition, shape = Department), size=3, alpha=.7) + 
  #geom_smooth(aes(col=Attrition), method="lm", se=F) +
  scale_color_manual(values = c("No" = "grey", "Yes" = "red")) +
  theme(panel.grid.major.y = element_line(color="grey"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_line(color="grey"),
        panel.grid.minor.y = element_line(color="grey"),
        panel.background = element_blank())


# Model Job Role + Job Satisfaction  ~ Attrition

ggplot(data, aes(x=JobLevel, y=Attrition)) +
  geom_jitter()


```


## 4. Predicting Attrition with knn   

  *  In this section, I create and evaluate several models to find a suitable model to apply to a dataset where Attrition is unknown  
  * Model 7 produces a model where both specificity and sensitivity are above 60%.  This model is applied to the Attrition Competition data set to predict attrition.  
  
```{r}
library(class)
library(caret)

# Load in the set that will be used as the final test set/competition set 
attrition_comp <- s3read_using(FUN = read.csv,
                     bucket = "ddsproject1",
                     object = "CaseStudy2CompSet No Attrition.csv")

# Change Attrition to numeric levels (Yes=1 / No=0)
data2 <- data %>%
  mutate(Attrition = recode(Attrition, Yes = "1", No = '0'))

# Create training indices for knn model  
set.seed(1)
splitPerc = .7

trainInd_attr = sample(1:dim(data2)[1], round(splitPerc * dim(data2)[1]))
train_attr = data2[trainInd_attr,]
test_attr = data2[-trainInd_attr,]

# Model 2: JobLevel and EnvironmentSatisfaction --------------------------------
train2 <- train_attr %>%
  select(JobLevel, EnvironmentSatisfaction, Attrition)

test2 <- test_attr %>%
  select(JobLevel,EnvironmentSatisfaction, Attrition)

knn_model2 = knn(train2[,c(1,2)], test2[,c(1,2)], train2$Attrition, k=5, prob=TRUE)
confusionMatrix(table(knn_model2, test2$Attrition)) 

# Change Threshold
probs = ifelse(knn_model2 == "1", attributes(knn_model2)$prob, 1- attributes(knn_model2)$prob) 

# New Threshold
summary(data2$Attrition)
# 140/870 = 16% left 

NewClass = ifelse(probs > .14, "1", "0")
table(NewClass, test2[,3])
CM = confusionMatrix(table(NewClass, test2[,3]), mode = "everything")
CM



# Model 3: JobInvolvement and RelationshipSatisfaction -------------------------
train3 <- train_attr %>%
  select(JobInvolvement, RelationshipSatisfaction, Attrition)

test3 <- test_attr %>%
  select(JobInvolvement, RelationshipSatisfaction, Attrition)

knn_model3 = knn(train3[,c(1,2)], test3[,c(1,2)], train3$Attrition, k=5, prob=TRUE)

confusionMatrix(table(knn_model3, test3$Attrition)) 

# Change threshold 
probs3 = ifelse(knn_model3 == "1", attributes(knn_model3)$prob, 1- attributes(knn_model3)$prob) 

#New Threshold
summary(data2$Attrition)
# 140/870 = 16% left 

NewClass3 = ifelse(probs3 > .17, "1", "0")
table(NewClass3, test3[,3])
CM3 = confusionMatrix(table(NewClass3, test3[,3]), mode = "everything")
CM3



# Model 4: JobInvolvement and Age ----------------------------------------------
train4 <- train_attr %>%
  select(JobInvolvement, Age, Attrition)

test4 <- test_attr %>%
  select(JobInvolvement, Age, Attrition)

knn_model4 = knn(train4[,c(1,2)], test4[,c(1,2)], train4$Attrition, k=5, prob=TRUE)

confusionMatrix(table(knn_model4, test4$Attrition)) 

# Change threshold 
probs4 = ifelse(knn_model4 == "1", attributes(knn_model4)$prob, 1- attributes(knn_model4)$prob) 

#New Threshold
summary(data2$Attrition)
# 140/870 = 16% left 

NewClass4 = ifelse(probs4 > .1, "1", "0")
table(NewClass4, test4[,3])
CM4 = confusionMatrix(table(NewClass4, test4[,3]), mode = "everything")
CM4


# Model 5: JobLevel and NumCompaniesWorked -------------------------------------
train5 <- train_attr %>%
  select(JobLevel, NumCompaniesWorked, Attrition)

test5 <- test_attr %>%
  select(JobLevel, NumCompaniesWorked, Attrition)

knn_model5 = knn(train5[,c(1,2)], test5[,c(1,2)], train5$Attrition, k=5, prob=TRUE)

confusionMatrix(table(knn_model5, test5$Attrition)) 

# Change threshold  
probs5 = ifelse(knn_model5 == "1", attributes(knn_model5)$prob, 1- attributes(knn_model5)$prob) 

#New Threshold
summary(data2$Attrition)
# 140/870 = 16% left 

NewClass5 = ifelse(probs5 > .16, "1", "0")
table(NewClass5, test5[,3])
CM5= confusionMatrix(table(NewClass5, test5[,3]), mode = "everything")
CM5


# Model 6: RelationshipSatisfaction and NumCompaniesWorked ---------------------
train6 <- train_attr %>%
  select(RelationshipSatisfaction, NumCompaniesWorked, Attrition)

test6 <- test_attr %>%
  select(RelationshipSatisfaction, NumCompaniesWorked, Attrition)

knn_model6 = knn(train6[,c(1,2)], test6[,c(1,2)], train6$Attrition, k=5, prob=TRUE)

confusionMatrix(table(knn_model6, test6$Attrition)) 

# Change threshold  
probs6 = ifelse(knn_model6 == "1", attributes(knn_model6)$prob, 1- attributes(knn_model6)$prob) 

#New Threshold
summary(data2$Attrition)
# 140/870 = 16% left 

NewClass6 = ifelse(probs6 > .18, "1", "0")
table(NewClass6, test6[,3])
CM6= confusionMatrix(table(NewClass6, test6[,3]), mode = "everything")
CM6

# Model 7: YearsInCurrent Role, YearsSinceLastPromotion, Age, JobLevel ---------
# Use this model - specificity and sensitivity are both over 60 ----------------

train7 <- train_attr %>%
  select(YearsInCurrentRole, YearsSinceLastPromotion, Age, JobLevel, Attrition)

test7 <- test_attr %>%
  select(YearsInCurrentRole, YearsSinceLastPromotion, Age, JobLevel, Attrition)

knn_model7 = knn(train7[,c(1,4)], test7[,c(1,4)], train7$Attrition, k=5, prob=TRUE)

confusionMatrix(table(knn_model7, test7$Attrition)) 

# Change threshold 
#Get probs of a FRAUD specifically
probs7 = ifelse(knn_model7 == "1", attributes(knn_model7)$prob, 1- attributes(knn_model7)$prob) 

#New Threshold
summary(data2$Attrition)
# 140/870 = 16% left 

NewClass7 = ifelse(probs7 > .17, "1", "0")
table(NewClass7, test7[,5])
CM7= confusionMatrix(table(NewClass7, test7[,5]), mode = "everything")
CM7


# Apply Model 7 to Attrition Competition Set -----------------------------------
comp7 <- attrition_comp %>%
  select(YearsInCurrentRole, YearsSinceLastPromotion, Age, JobLevel)

knn_comp7 = knn(train7[,c(1,3)], comp7[,c(1,3)], train7$Attrition, k=5, prob=TRUE)
knn_comp7

summary(knn_comp7)

write.csv(attrition_comp,"C:\\Users\\48103982\\Documents\\MSDS\\6306 Doing Data Science\\Project 2\\CaseStudy2PredictionsTiczon_Attrition.csv")

```

## 5. Predicting Attrition with Linear Regression  

  * Here I attempt to create a model using linear regression; I do not ultimately use this for prediction  
  
```{r}
data3 <- data2 # Create copy of data frame where some variables are numeric,  not factor level

data3$JobLevel = as.numeric(data3$JobLevel) 
data3$EnvironmentSatisfaction = as.numeric(data3$EnvironmentSatisfaction) 
data3$Attrition = as.numeric(as.character(data3$Attrition))

lm_model1 <- lm(Attrition ~ JobLevel + EnvironmentSatisfaction, data=data3)
print(lm_model1)
summary(lm_model1)

predict(lm_model1)  
#predict(lm_model1, newdata= comp2) 
```

## 6. Predicting Attrition with Naive Bayes   

 * Here I attempt to create a model using Naive Bayes; I do not ultimately use this for prediction  
```{r}
#install.packages("e1071")
#update.packages()
Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS=TRUE) # run this if e1071 isn't loading 
library(e1071) 


NBtrainIndices = sample(seq(1:length(data3$EnvironmentSatisfaction)),round(.7*length(data3$EnvironmentSatisfaction)))
NBtrain = data3[NBtrainIndices,]
NBtest = data3[-NBtrainIndices,]

nb_model = naiveBayes(NBtrain[,c("EnvironmentSatisfaction","JobLevel")], factor(NBtrain$Attrition, labels = c("No", "Yes")))

nbCM = confusionMatrix(table(factor(NBtest$Attrition, 
                                  labels = c("No", "Yes")),
                           predict(nb_model, NBtest[,c("EnvironmentSatisfaction","JobLevel")])))
nbCM
```



## 7. Predicting Salary  

  *  In this section, I create and evaluate a linear regression model to predict Monthly Income of employees in the Salary Competition data set
  *  The model uses Job Level as the predictor variable  
  *  The model's RMSE is \$1414.96 which is less than the maximum RMSE (\$3000) requirement of the model 
```{r}
salary_comp <- s3read_using(FUN = read.csv,
                     bucket = "ddsproject1",
                     object = "CaseStudy2CompSet No Salary.csv")


# Create and evaluate model ----------------------------------------------------

pred_error_sq <- c(0)
for(i in 1:dim(data3)[1]) {
 salary_train <- data3[-i,]
  fit <- lm(MonthlyIncome ~ JobLevel, data = salary_train)  
  salary_i <- predict(fit, data.frame(JobLevel = data3[i,16]))  
  pred_error_sq <- pred_error_sq + (data3[i,20] - salary_i)^2  
}

SSE = var(salary_comp$JobLevel) * (869)

R_squared <- 1 - (pred_error_sq/SSE) 
R_squared

MSE = pred_error_sq / 870
MSE

RMSE = sqrt(pred_error_sq/870)
RMSE # $1414.96 < $3000 

# Apply model to Salary Competition Set ----------------------------------------

predicted_monthly_income <- numeric(dim(salary_comp)[1])  # Initialize a vector to store predictions

for(i in 1:dim(salary_comp)[1]) {
  salary_train <- data3  # Use original dataset for training the model
  fit <- lm(MonthlyIncome ~ JobLevel, data = salary_train)
  salary_i <- predict(fit, newdata = data.frame(JobLevel = salary_comp[i, "JobLevel"]))
  predicted_monthly_income[i] <- salary_i
}
salary_comp$PredictedMonthlyIncome <- predicted_monthly_income

write.csv(salary_comp, "C:\\Users\\48103982\\Documents\\MSDS\\6306 Doing Data Science\\Project 2\\CaseStudy2PredictionsTiczon_Salary.csv")

```

